PRIMUS: Galaxy Environment on the Quiescent Fraction Evolution at $z < 1$

### 2014-08-21 qfenv_paper/qf_environment.tex
Mike's revision part 1 finished

\begin{table} %Subject to significant change based on the classification of environment
  \caption{Galaxy Subsamples}
  \label{tab:subsample}
  \begin{center}
    \leavevmode
    \begin{tabular}{l l l l l l l } \hline \hline              
    $z$ &$n_{\rm{env}}$        &Quiescent  &Star-forming  & $\mathcal{M}_{\rm{lim}}$ & $M_{\rm{r}}$ cut-off \\ \hline 
$0.05-0.12$ &Low           &12859                       &14443                           \\
               &High            &18952                       &13387                           \\ 
                              &               &                       &                           \\ \hline
$0.2-0.4$      &Low           &1158                    &3445                           \\
               &High            &817                    &1531                           \\
               &               &                       &                           \\ \hline
$0.4-0.6$      &Low           &1315                       &3390                           \\
               &High            &848                       &1499                           \\
               &               &                       &                           \\ \hline
$0.6-0.8$      &Low           &1158                       &2303                           \\
               &High            &900                       &1275                           \\
               &               &                       &                           \\ \hline
  \multicolumn{4}{l}{}                                             \\       
    \end{tabular} \par
    \end{center}
%    \bigskip 
    {\bf Notes}: Number of galaxies in the mass complete subsamples derived in Section \ref{sec:sample}. The subsamples are classified based on environment $n_{\rm{env}}$ and star formation rate (star-forming or quiescent). The lowest redshift bin is derived from SDSS; the rest are from PRIMUS. 
    \bigskip
\end{table}

@@

We present the absolute magnitude ($M_{r}$) versus redshift for the galaxy sample (black squares) in Figure \ref{fig:targetEDP}. The left-most panel corresponds to the sample derived from the SDSS-{\em GALEX} data and the rest correspond to the target sample derived from the PRIMUS data divided in bins with $\Delta z \sim 0.2$. 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% CLASSIFYING QUIESCENT AND STAR-FORMING GALAXIES
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Classifying Quiescent and Star-Forming Galaxies} \label{sec:sfq}
We now classify our mass complete galaxy sample into quiescent or star-forming using an evolving cut based on specific star-formation rate utilized in \cite{Moustakas:2013aa} Section 3.2. This classification method uses the star-forming (SF) sequence, which is the correlation between star-formation rate (SFR) and stellar mass in star-forming galaxies observed at least until $z \sim 2$ (\citealt{Noeske:2007aa}). The PRIMUS sample displays a well-defined SF sequence within the redshift range of our galaxy sample. Using the power-law slope for the SF sequence from \cite{Salim:2007aa} (SFR $\propto \mathcal{M}^{0.65}$) and the minimum of the quiescent/star-forming bimodality, determined empirically, we obtain the following equation to classify the target galaxies (Equation 2 in \citealt{Moustakas:2013aa}):
\begin{equation}
\rm{log}(\rm{SFR}_{\rm{min}}) = -0.49 + 0.64 \rm{log}(\mathcal{M} - 10) +1.07(z-0.1), 
\end{equation} \label{eq:qsfclass} 
where $\mathcal{M}$ is the stellar mass of the galaxy. If the target galaxy SFR and stellar mass lies above Equation \ref{eq:qsfclass} we classify it as star-forming; if below, as quiescent (\cite{Moustakas:2013aa} Figure 1.).

\begin{figure*}
  \begin{center}
    \leavevmode
    \epsfig{file=fig_smf_cylr2h25_thresh75_bin0_20_sdss_z0_05_0_12_primus_z0_2_1_0_2envbin_lowenv0_5_highenv3_0_lit_primuszerr.eps, height=0.75\textwidth,angle=90}
     \caption{Evolution of stellar mass functions of star-forming (top) and quiescent (bottom) target galaxies in 
low (left) and high (right) environments from redshift range $z=0-0.8$. The environment of each galaxy  
was calculated using a cylindrical aperture size of $R=2 \: \rm{Mpc}$ and $H=25 \: \rm{Mpc}$ and  
classification based on the cut-offs specified in Table \ref{tab:aperture}. The SMFs use mass bins of 
width $\Delta \rm{log}(\mathcal{M}/\mathcal{M}_{\odot})=0.25$. In each panel we use shades of blue 
(star-forming) and orange (quiescent) to represent the SMF at different redshift, higher redshifts being
progressively lighter.}      \label{fig:smf}
    \end{center}
\end{figure*}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% GALAXY ENVIRONMENT
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Galaxy Environment} \label{sec:environment}
We define the environment of a galaxy as the number of neighboring {\em Environment Defining Population} galaxies (defined below) within a fixed aperture centered around it. We use a fixed aperture measurement of environment because, as \cite{Muldrew:2012aa} finds in their comparison of different environment definition using simulations, it provides a better probe of the entire halo in comparison to other definitions of environment such as nearest neighbor, which provides a better tracer at inter-halo scales.

For our aperture, we use a cylinder of dimensions: $R_{\rm{ap}} = 2 \; \rm{Mpc}/h$ and $H_{\rm{ap}} = 25 \; \rm{Mpc}/h$. Though spherical apertures are often used in literature (e.g. \cite{Croton:2005aa}), we use a cylindrical aperture in order to account for the PRIMUS redshift errors and redshift space distortions (i.e. "Finger of God" effect). \cite{Cooper:2005aa} finds that $\pm 1000 \; \rm{km} \; \rm{s^{-1}}$ optimally reduces the effects of redshift space distortions. The PRIMUS redshift uncertainty at $z \sim 0.7$ corresponds to $\sigma_z < 0.01$, so our aperture height choice of $25 \; \rm{Mpc}/h$ accounts for both of these effects. Our choice of cylinder radius was motivated by scale dependence analyses in literature (\cite{Blanton:2006aa}, \cite{Wilman:2010aa}, and \cite{Muldrew:2012aa}), which suggest that galactic properties such as color and quenched fractions are dependent on halo-scale properties such as host dark matter halo mass. \cite{Wilman:2010aa}, which uses environment defined by annuli of different radii, find positive correlation for quenched fraction and color on scales $< 1 \; \rm{Mpc}$ and anti-correlation on scales $> 3 \; \rm{Mpc}$. Our choice of $2 \rm{Mpc}/h$ provides sufficient sample size of galaxies in dense environments, for robust statistics, while tracing galactic properties within the halo scale.

@@

Before we measure the environment for our galaxy sample, we first construct a volume limited {\em Environment Defining Population} (EDP) with absolute magnitude cut-offs ($M_{r}$) at redshift bin of $\Delta z \sim 0.2$. The $M_{r}$ cut-offs for the EDP are selected such that the cumulative number density over $M_{r}$ for all redshift bins are equal. In addition to providing a reasonable comparison of environment measurements for wide redshift range, it also seeks to construct an EDP that contains similar galaxy populations through the redshift range (i.e. accounts for the progenitor bias). As \cite{Behroozi:2013aa} and \cite{Leja:2013aa} find in their analysis of the cumulative number density method, though it does not precisely account for the scatter in mass accretion or galaxy-galaxy mergers, it provides a reasonable means to compare galaxy populations over a wide range of cosmic time.

@@

For the EDP of the SDSS-{\em GALEX} galaxy sample (hereafter SDSS EDP), we do not use the SDSS-{\em GALEX} parent data which uses the geometry of the combined angular selection function of the VAGC and {\em GALEX} (Section \ref{sec:sdss}). Instead, since FUV, NUV values are not necessary for EDP, we extend the parent data of the SDSS EDP to the entire VAGC, including galaxies outside of the {\em GALEX} window function. Furthermore, we impose a redshift range of $0.05-0.12$ on the SDSS EDP. This redshift range was empirically determined to account for the lack of faint galaxies at $z \sim 0.2$ and the lack of bright galaxies at $z \sim 0.01$ in the VAGC. The lower bound was determined by the bright limit and the upper bound by the faint limit of the $M_r$ versus redshift distribution. As with the PRIMUS redshift bins, we determine the SDSS EDP $M_{r, \rm{lim}}$ by matching the cumulative number density of the highest redshift bin. For redshift bins $0.05-0.12$, $0.2-0.4$, $0.4-0.6$, $0.6-0.8$ we get $M_{r,\rm{lim}} = -20.57$, $-20.73$, $-20.80$ and $-20.95$, respectively. These absolute magnitude limits are illustrated in Figure \ref{fig:targetEDP}, which show clear $M_r$ cutoffs in the $M_{r}$ distribution versus redshift for the EDP (red circles) on top of the target galaxy sample (black).

For our SDSS-{\em GALEX} galaxy sample, in Section \ref{sec:target}, we imposed PRIMUS redshift errors in order to establish a consistent measurement of environment throughout our redshift range. We appropriately impose equivalent redshift adjustments for the SDSS EDP. So for the SDSS EDP galaxies that are within the SDSS-{\em GALEX} sample, we adjust the redshift by an identical amount. For the rest, we apply the same redshift adjustment procedure described in Section \ref{sec:target} in order to obtain PRIMUS level redshift uncertainties.

Finally we measure the environment for each galaxy in our galaxy sample by counting the number of EDP galaxies, $n_{\rm{env}}$ with $RA$, $Dec$, and $z$ within our cylindrical aperture centered around it. $n_{\rm{env}}$ accounts for the statistical weights of the EDP galaxies. Once we obtain environment measurements for all the galaxies in our galaxy sample, we classify galaxies with $n_{\rm{env}} < 1$ to be "low" environment density and galaxies with $n_{\rm{env}} > 3$ to be "high" environment densities. The high environment cutoff was selected in order to reduce contamination from galaxies in low environment densities while maintaining sufficient statistics.

The analysis we describe below uses a fixed cylindrical aperture with dimensions $R_{\rm{ap}} = 2 \: \rm{Mpc}$ and $H_{\rm{ap}} = 25 \: \rm{Mpc}$ to measure environment. The same analysis was extended for varying aperture dimensions $R_{\rm{ap}} = 1, 2, 3 \: \rm{Mpc}$ and $H_{\rm{ap}} = 25, 50 \: \rm{Mpc}$. We also adjusted the environment classification thresholds for these analyses. The results obtained from using different apertures and environment classification thresholds are consistent with the results later presented.

@@

To account for these edge effects, we use a Monte Carlo method to impose edge cutoffs on our galaxy sample. We first construct a \texttt{ransack} sample of  $N_{\rm{ransack}} = 1,000,000$ points with $RA$ and $Dec$ randomly selected within the window function of the EDP (so SDSS EDP and PRIMUS EDP separately). We then compute the angular separation, $\theta_{i, \rm{ap}}$ that corresponds to $R_{\rm{ap}}$ (Section \ref{sec:environment}) at the redshift of each sample galaxy $i$. For each sample galaxy we count the number of \texttt{ransack} points within $\theta_{i, \rm{ap}}$ of the galaxy: $n_{i,\rm{ransack}}$. Afterwards, we compare $n_{i,\rm{ransack}}$ to the expected value computed from the angular area of the environment defining aperture and the EDP window function: 
\begin{equation} \label{eq:ransack}
E_{i}[n_{\rm{ransack}}] = \frac{N_{\rm{ransack}}}{A_{\rm{EDP}}}\times {\pi \theta_{i, \rm{ap}}^2} \times f_{\rm{thresh}}. 
\end{equation} 
$A_{\rm{EDP}}$ is the total angular area of the EDP window function and $f_{\rm{thresh}}$ is the fractional threshold for the edge effect cut-off. For $R_{\rm{ap}}= 2\;\rm{Mpc}$, $f_{\rm{thresh}} = 0.75$. If $n_{i, \rm{ransack}} > E_{i}[n_{\rm{ransack}}]$ then galaxy $i$ remains in our sample; otherwise, it is discarded.

>>>>

\begin{table} %Subject to significant change based on the classification of environment
  \caption{Galaxy Subsamples}
  \label{tab:subsample}
  \begin{center}
    \leavevmode
    \begin{tabular}{l l l l l l l } \hline \hline              
    $z$ &$n_{\rm{env}}$        &Quiescent  &Star-forming  & $\mathcal{M}_{\rm{lim}}$ & $M_{\rm{r}, lim}$ \\ \hline 
$0.05-0.12$ &Low           &12859                       &14443                           \\
               &High            &18952                       &13387                           \\ 
                              &               &                       &                           \\ \hline
$0.2-0.4$      &Low           &1158                    &3445                           \\
               &High            &817                    &1531                           \\
               &               &                       &                           \\ \hline
$0.4-0.6$      &Low           &1315                       &3390                           \\
               &High            &848                       &1499                           \\
               &               &                       &                           \\ \hline
$0.6-0.8$      &Low           &1158                       &2303                           \\
               &High            &900                       &1275                           \\
               &               &                       &                           \\ \hline
  \multicolumn{4}{l}{}                                             \\       
    \end{tabular} \par
    \end{center}
%    \bigskip 
    {\bf Notes}: Number of galaxies in the mass complete subsamples derived in Section \ref{sec:sample}. The subsamples are classified based on environment $n_{\rm{env}}$ and star formation rate (star-forming or quiescent). The lowest redshift bin is derived from SDSS; the rest are from PRIMUS. 
    \bigskip
\end{table}

@@

We present the absolute magnitude ($M_{r}$) versus redshift for the galaxy sample (black squares) in Figure \ref{fig:targetEDP}. The left-most panel corresponds to the sample derived from the SDSS-{\em GALEX} data and the rest correspond to the target sample derived from the PRIMUS data divided in bins with $\Delta z \sim 0.2$. 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% CLASSIFYING QUIESCENT AND STAR-FORMING GALAXIES
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Classifying Quiescent and Star-Forming Galaxies} \label{sec:sfq}
We now classify our mass complete galaxy sample into quiescent or star-forming using an evolving cut based on specific star-formation rate utilized in \cite{Moustakas:2013aa} Section 3.2. This classification method uses the star-forming (SF) sequence, which is the correlation between star-formation rate (SFR) and stellar mass in star-forming galaxies observed at least until $z \sim 2$ (\citealt{Noeske:2007aa}). The PRIMUS sample displays a well-defined SF sequence within the redshift range of our galaxy sample. Using the power-law slope for the SF sequence from \cite{Salim:2007aa} (SFR $\propto \mathcal{M}^{0.65}$) and the minimum of the quiescent/star-forming bimodality, determined empirically, we obtain the following equation to classify the target galaxies (Equation 2 in \citealt{Moustakas:2013aa}):
\begin{equation}
\rm{log}(\rm{SFR}_{\rm{min}}) = -0.49 + 0.64 \rm{log}(\mathcal{M} - 10) +1.07(z-0.1), 
\end{equation} \label{eq:qsfclass} 
where $\mathcal{M}$ is the stellar mass of the galaxy. If the target galaxy SFR and stellar mass lie above Equation \ref{eq:qsfclass} we classify it as star-forming; if below, as quiescent (\citealt{Moustakas:2013aa} Figure 1.).

\begin{figure*}
  \begin{center}
    \leavevmode
    \epsfig{file=fig_smf_cylr2h25_thresh75_bin0_20_sdss_z0_05_0_12_primus_z0_2_1_0_2envbin_lowenv0_5_highenv3_0_lit_primuszerr.eps, height=0.75\textwidth,angle=90}
     \caption{Evolution of stellar mass functions of star-forming (top) and quiescent (bottom) target galaxies in 
low (left) and high (right) environments from redshift range $z=0-0.8$. The environment of each galaxy  
was calculated using a cylindrical aperture size of $R=2 \: \rm{Mpc}$ and $H=25 \: \rm{Mpc}$ and  
classification based on the cut-offs specified in Table \ref{tab:aperture}. The SMFs use mass bins of 
width $\Delta \rm{log}(\mathcal{M}/\mathcal{M}_{\odot})=0.25$. In each panel we use shades of blue 
(star-forming) and orange (quiescent) to represent the SMF at different redshift, higher redshifts being
progressively lighter.}      \label{fig:smf}
    \end{center}
\end{figure*}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% GALAXY ENVIRONMENT
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Galaxy Environment} \label{sec:environment}
We define the environment of a galaxy as the number of neighboring Environment Defining Population galaxies (defined below) within a fixed aperture centered around it. We use a fixed aperture measurement of environment in order to probe environment on the halo scale, as \cite{Muldrew:2012aa} finds in their comparison of different environment definition using simulations.

For our aperture, we use a cylinder of dimensions: $R_{\rm{ap}} = 2 \; h^{-1} \rm{Mpc}$ and $H_{\rm{ap}} = 25 \; h^{-1}\rm{Mpc}$. Though spherical apertures are often used in literature (e.g. \citealt{Croton:2005aa}), we use a cylindrical aperture in order to account for the PRIMUS redshift errors and redshift space distortions (i.e. ``Finger of God" effect). \cite{Cooper:2005aa} finds that $\pm 1000 \; \rm{km} \; \rm{s^{-1}}$ optimally reduces the effects of redshift space distortions. The PRIMUS redshift uncertainty at $z \sim 0.7$ corresponds to $\sigma_z < 0.01$, so our aperture height choice of $25 \; \rm{Mpc}/h$ accounts for both of these effects. Our choice of cylinder radius was motivated by scale dependence analyses in literature (\citealt{Blanton:2006aa, Wilman:2010aa, Muldrew:2012aa}), which suggest that galactic properties such as color and quenched fractions are dependent on halo-scale properties such as host dark matter halo mass. \cite{Wilman:2010aa}, which uses environment defined by annuli of different radii, find positive correlation for quenched fraction and color on scales $< 1 \; \rm{Mpc}$ and anti-correlation on scales $> 3 \; \rm{Mpc}$. Our choice of $2 \rm{Mpc}/h$ provides sufficient sample size of galaxies in dense environments, for robust statistics, while tracing galactic properties within the halo scale.

@@

Before we measure the environment for our galaxy sample, we first construct a volume limited Environment Defining Population (EDP) with absolute magnitude cut-offs ($M_{r}$) at redshift bin of $\Delta z \sim 0.2$. The $M_{r}$ cut-offs for the EDP are selected such that the cumulative number density over $M_{r}$ for all redshift bins are equal. In addition to providing a reasonable comparison of environment measurements for wide redshift range, we make this choice in order to construct an EDP that contains similar galaxy populations through the redshift range (i.e. accounts for the progenitor bias). As \cite{Behroozi:2013aa} and \cite{Leja:2013aa} find in their analysis of the cumulative number density method, though it does not precisely account for the scatter in mass accretion or galaxy-galaxy mergers, it provides a reasonable means to compare galaxy populations over a wide range of cosmic time.

@@

For the EDP of the SDSS-{\em GALEX} galaxy sample (hereafter SDSS EDP), we do not use the SDSS-{\em GALEX} parent data, which uses the geometry of the combined angular selection function of the VAGC and {\em GALEX} (Section \ref{sec:sdss}). Instead, since FUV, NUV values are not necessary for the EDP, we extend the parent data of the SDSS EDP to the entire NYU VAGC, including galaxies outside of the {\em GALEX} window function. Furthermore, we impose a redshift range of $0.05-0.12$ on the SDSS EDP. This redshift range was determined to account for the lack of faint galaxies at $z \sim 0.2$ and the lack of bright galaxies at $z \sim 0.01$ in the VAGC. As with the PRIMUS redshift bins, we determine the SDSS EDP $M_{r, \rm{lim}}$ by matching the cumulative number density of the highest redshift bin. For redshift bins $0.05-0.12$, $0.2-0.4$, $0.4-0.6$, $0.6-0.8$ we get $M_{r,\rm{lim}} = -20.57$, $-20.73$, $-20.80$ and $-20.95$, respectively. These absolute magnitude limits are illustrated in Figure \ref{fig:targetEDP}, which show clear $M_r$ cutoffs in the $M_{r}$ distribution versus redshift for the EDP (red circles) on top of the target galaxy sample (black).

For our SDSS-{\em GALEX} galaxy sample, in Section \ref{sec:target}, we apply PRIMUS redshift errors in order to establish a consistent measurement of environment throughout our redshift range. We appropriately apply equivalent redshift adjustments for the SDSS EDP. So for the SDSS EDP galaxies that are within the SDSS-{\em GALEX} sample, we adjust the redshift by an identical amount. For the rest, we apply the same redshift adjustment procedure described in Section \ref{sec:target} in order to obtain PRIMUS level redshift uncertainties.

Finally we measure the environment for each galaxy in our galaxy sample by counting the number of EDP galaxies, $n_{\rm{env}}$ with $RA$, $Dec$, and $z$ within our cylindrical aperture centered around it. $n_{\rm{env}}$ accounts for the statistical weights of the EDP galaxies. Once we obtain environment measurements for all the galaxies in our galaxy sample, we classify galaxies with $n_{\rm{env}} < 0.5$ to be ``low" environment density and galaxies with $n_{\rm{env}} > 3$ to be ``high" environment densities. The high environment cutoff was selected in order to reduce contamination from galaxies in low environment densities while maintaining sufficient statistics.

The analysis we describe below uses a fixed cylindrical aperture with dimensions $R_{\rm{ap}} = 2 \: \rm{Mpc}$ and $H_{\rm{ap}} = 25 \: \rm{Mpc}$ to measure environment. The same analysis was extended for varying aperture dimensions $R_{\rm{ap}} = 1, 2, 3 \: \rm{Mpc}$ and $H_{\rm{ap}} = 25, 50 \: \rm{Mpc}$. We also adjusted the environment classification thresholds for these analyses. The results obtained from using different apertures and environment classification thresholds are consistent with the results presented below.

@@

To account for these edge effects, we use a Monte Carlo method to impose edge cutoffs on our galaxy sample. First, using \texttt{ransack} from \cite{Swanson:2008aa}, we construct a random sample of  $N_{\rm{ransack}} = 1,000,000$ points with $RA$ and $Dec$ randomly selected within the window function of the EDP (SDSS EDP and PRIMUS EDP separately). We then compute the angular separation, $\theta_{i, \rm{ap}}$ that corresponds to $R_{\rm{ap}}$ (Section \ref{sec:environment}) at the redshift of each sample galaxy $i$. For each sample galaxy we count the number of \texttt{ransack} points within $\theta_{i, \rm{ap}}$ of the galaxy: $n_{i,\rm{ransack}}$. Afterwards, we compare $n_{i,\rm{ransack}}$ to the expected value computed from the angular area of the environment defining aperture and the EDP window function: 
\begin{equation} \label{eq:ransack}
<n_{\rm{ransack}}>_{i} = \frac{N_{\rm{ransack}}}{A_{\rm{EDP}}}\times {\pi \theta_{i, \rm{ap}}^2} \times f_{\rm{thresh}}. 
\end{equation} 
$A_{\rm{EDP}}$ is the total angular area of the EDP window function and $f_{\rm{thresh}}$ is the fractional threshold for the edge effect cut-off. For $R_{\rm{ap}}= 2\;\rm{Mpc}$, $f_{\rm{thresh}} = 0.75$. If $n_{i, \rm{ransack}} > E_{i}[n_{\rm{ransack}}]$ then galaxy $i$ remains in our sample; otherwise, it is discarded.
